Data Management Plan

### 2021-11-29 application/dataManagementPlan_long_joaco.tex
Several modifications wrt dataManagementPlan.tex

This is a software project.  We will produce and manage new modules to be integrated into the Bonsai software ecosystem.  These will include communication modules to be written in C\#, and machine-learning applications written in Python, R and Matlab.

@@

\subsection{Managing, storing and curating data}

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at
\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
contribute to this forum (Figure~1f, \emph{Case for Support}).

@@

\textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term. Specify any community agreed or other formal data standards used (with URL references).}}

Bonsai is a freely available open-source resource, already widely adopted by
the systems neuroscience and behavioural science communities (see Fig 1f).
Current contributors include other large open projects (OpenEphys, DeepLabCut),
and companies developing and supplying new hardware devices
(Neurophotometrics). This community uptake, engagement, and commitment provides
a foundation for long-term sustainability.

NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
current proposal. Founded by the original developer of Bonsai, it helps to
support and expand the user base well beyond the life of the current proposal.
NeuroGEARS currently provides quality control and curation for the core Bonsai
code, and provides training and support services for research institutions
around the world who want to start using Bonsai. NeuroGEARS also facilitates
knowledge transfer between users and institutions adopting Bonsai, by
moderating the community user forums and collaborating with organisations who
want to build new innovative research projects on top of Bonsai, as in the
current proposal.

The specific contributions developed as part of this project will form new core
components and packages within the Bonsai ecosystem. In particular, the new
language integration packages will support long-term sustainability, by
dramatically increasing Bonsai's interoperability with existing software
development communities which are already well established in basic science
(MATLAB, Python, R). Developing these new packages with input from the core
advisory panel of Bonsai, and early public distribution and user testing within
the SWC and with planned workshops, will ensure solid and broad user engagement
from the beginning of the development cycle and will therefore build momentum
for their use, support and dissemination. Close collaboration with the core
maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
responsibility for maintenance and support of these packages beyond the end of
the project, following the same principles of open-source and free sharing of
knowledge that characterises the entire Bonsai ecosystem.

@@

\textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

The fact that Bonsai is distributed as an open source project
(\url{https://bonsai-rx.org/}), that is has extensive online documentation
(\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
(\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
enables the use of the Bonsai ecosystem by the general public.

@@

Plans and place for long-term storage, preservation, and planned retention period for the research data. Formal preservation standards, if any. Indicate which data may not be retained (if any). Also, briefly describe what will be stored long-term, raw data, processed data?

\section{Data sharing and access}
\subsection{Where will data be shared?}
% Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
Software modules and documentation will be available on github, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.

@@

Release management: repositories will be public throughout, but marked as developmental.  Alpha-level code will be available but not publicised until after extensive testing.  Will be integrated into package management and advertised once it achieves release quality.

>>>>

This is a software project.  We will produce and manage new modules to be
integrated into the Bonsai software ecosystem.  These will include extensions
to incorporate non-native code packages into Bonsai (Section 2.1, Case for
Support), a general abstraction framework for data processing (Section 2.2,
Case for Support), machine learning functionality added to Bonsai (Section 2.3,
Case for Support) and documentation.

@@

\subsection{Managing, storing and curating data}
% \textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term. Specify any community agreed or other formal data standards used (with URL references).}}

% code
The source code for the core Bonsai ecosystem and the majority of extension
packages are hosted at \url{github.com/bonsai-rx} (MIT license with no
copyright and licencing restrictions).

@@

\subsubsection*{Team responsibilities}

\begin{description}

    \item[Goncalo Lopes] Bonsai related software development,
        software releases, organization of Bonsai workshops
        and co-supervision of research software engineer (RSE).

    \item[Joaquin Rapela] machine learning (ML) related software development,
        interaction with ML functionality methods developers, validating of ML
        functionality with behavioral and neural recordings and co-supervision
        of RSE.

    \item[Maneesh Sahani] advise on ML.

    \item[Tom Mrsic-Flogel] advise on neuroscience topics and communication
        with the steering committee.

\end{description}

\subsection*{Resourcing and responsibility}

\begin{description}

    \item[software development model] we will use the Kanban software
        development model, which we both Dr.~Lopes and Dr.~Rapela are already
        using in a collaborative project at the SWC/GCNU.

    \item[weekly meetings] between Goncalo Lopes, Joaquin Rapela and the RSE
        will be an essential component of the software development process.
        These will ensure that every member of the team is informed about both
        the Bonsai integration and the machine learning developments and will
        make the software development team more reliable upon the absence of
        any of its members.

    \item[code review and pair-programming] sessions will be held on a monthly basis among Goncalo Lopes, Joaquin Rapela and the recruited software engineer to deliver readable code sharing a standard programming style.

\end{description}

\subsubsection*{Major risks}

\begin{description}

    \item[language integration] we proposed to use existing software libraries
        to enable the communication between Bonsai, written in C\#, and other
        machine learning algorithms, written in Python (Python.NET), R (R.Net)
        and Matlab (native Matlab .Net interface). We have not yet tested any
        of these existing libraries, and we may found that they don't function
        reliably enough for our purposes. If this were the case, we proposed to
        build our own interface protocol using the ZeroMQ messaging library.
        This risk will only delay the progress of the proposed project, but
        will not hinder the deployment of its proposed outputs.

    \item[disfunctional machine learning functionality] in Section 2.3 of the
        Case for Support document (Figure 2, green box) we proposed to
        integrate existing machine learning algorithms into the Bonsai
        ecsosystem. It could be the case that some of these algorithms do not
        work as reliably as needed for our resource. I such a case we will use
        another algorithm with similar functionality.  This risk should not be
        a major problem, since most of the machine learning functionality we
        proposed to incorporate into Bonsai has been developed by us, or we are
        experienced using it.

    \item[lack of interest from Bonsai users] We will minimize the probability
        of this problem by advertising broadly the new functionality upon
        release.  However, we don't think this would be a real problem, since
        Bonsai users have reapeatly asked the addition of machine learning
        functionality to Bonsai (Section 3 Community demand, Case for Support),
        and because the machine learning functionality we proposed to add to
        Bonsai are popular among experimental-control neurocientists.

    \item[lack of interest from machine-learning methods developers] To
        minimize the probability that this risk turns into a large problem we
        will take three actions. First, we will provide machine-learning
        methods developers and easy-to-use API to interface their methods into
        the Bonsai ecosystem. Second, we will show machine-learning methods
        developers the large user exposure that Bonsai provides to integrated
        ML functionality. Third, we will organize Bonsai workshops for methods
        developers showcasing machine learning functionality integrated into
        the Bonsai ecosystem and instructing them how to integrate new methods
        into Bonsai.

    \end{description}

% Bonsai is a freely available open-source resource, already widely adopted by the systems neuroscience and behavioural science communities (see Fig 1f).  Current contributors include other large open projects (OpenEphys, DeepLabCut), and companies developing and supplying new hardware devices (Neurophotometrics). This community uptake, engagement, and commitment provides a foundation for long-term sustainability.

% NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the current proposal. Founded by the original developer of Bonsai, it helps to support and expand the user base well beyond the life of the current proposal.  NeuroGEARS currently provides quality control and curation for the core Bonsai code, and provides training and support services for research institutions around the world who want to start using Bonsai. NeuroGEARS also facilitates knowledge transfer between users and institutions adopting Bonsai, by moderating the community user forums and collaborating with organisations who want to build new innovative research projects on top of Bonsai, as in the current proposal.

% The specific contributions developed as part of this project will form new core components and packages within the Bonsai ecosystem. In particular, the new language integration packages will support long-term sustainability, by dramatically increasing Bonsai's interoperability with existing software development communities which are already well established in basic science (MATLAB, Python, R). Developing these new packages with input from the core advisory panel of Bonsai, and early public distribution and user testing within the SWC and with planned workshops, will ensure solid and broad user engagement from the beginning of the development cycle and will therefore build momentum for their use, support and dissemination. Close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over responsibility for maintenance and support of these packages beyond the end of the project, following the same principles of open-source and free sharing of knowledge that characterises the entire Bonsai ecosystem.

@@

% \textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

% The fact that Bonsai is distributed as an open source project
% (\url{https://bonsai-rx.org/}), that is has extensive online documentation
% (\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
% (\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
% many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
% enables the use of the Bonsai ecosystem by the general public.

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at the public
repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
700 users contribute to this forum (Figure~1f, \emph{Case for Support}).

Classes, methods and functions will be documented in the code, and automatic
documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
will be used to automatically generate API documentation. API documentation
will be complemented with class and interaction diagrams, documenting the
structure and dynamics of the application objects.

@@

\textcolor{orange}{\textit{Plans and place for long-term storage, preservation,
and planned retention period for the research data. Formal preservation
standards, if any. Indicate which data may not be retained (if any). Also,
briefly describe what will be stored long-term, raw data, processed data?}}

For persistent and long term preservation we propose to use
Zenodo\footnote{\url{https://zenodo.org/}}.

\section{Data sharing and access}
\subsection{Where will data be shared?}
% Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.

Software modules and documentation will be available on Github, with links from
the Bonsai project site (https://bonsai-rx.org) and available through the
Bonsai system's package management interface.

@@

% Release management: repositories will be public throughout, but marked as developmental.  Alpha-level code will be available but not publicised until after extensive testing.  Will be integrated into package management and advertised once it achieves release quality.

Method to accept a project output: software will be automatically tested with a
comprehensive set of unit test and integration tests, implemented as  GitHub
Actions. In addition, prior to distribution to the general public, Bonsai users
inside the SWC will evaluate the functionality of the software with their own
state-of-the-art datasets. Project outputs will be accepted when all tests run
successfully and no major functional problems are reported by SWC Bonsai users.

Release management: we will create a new release after each of the machine
learning functionalities (described in Section~2.3 of the Case for Support) is
integrated into the Bonsai ecosystem following the above method to accept a
project output. Credentials for release will be granted initially by Dr.
Gonçalo Lopes, Head Engineer at Neurogears and the current curator of the
Bonsai official package feed, but might be expanded to other team members after
training.  Acceptance criteria for release will be a group decision between
Dr.~Lopes, Dr.~Rapela, Prof.~Sahani and Prof.~Mrsic-Flogel.
